---
title: "BANA 7052: Lecture 06"
subtitle: "Model Building and Selection"
author: "Brandon M. Greenwell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: [default, metropolis, metropolis-fonts, hygge, "custom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: clear, middle, center

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, servr.daemon = TRUE, 
        crayon.enabled = FALSE)

# Global chunk options
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE,
  dev = "svglite",
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  error = FALSE
)

# Bitmoji id
my_id <- "1551b314-5e8a-4477-aca2-088c05963111-v1"

# Load required packages
library(ggplot2)
```

.font300[

[R code for these slides](https://github.com/bgreenwell/uc-bana7052/blob/master/code/lecture-06.R)

]


---
class: clear, middle

.font150[

* Required reading

    - Chapters: 9

    - Sections: 9.3-9.4

* Main topics:

    - Model selection

]


---

# Prerequisites

.code100[

```{r prerequisites, eval=FALSE}
# List of required (CRAN) packages
pkgs <- c(
  "car",          # for subsets() function
  "ggplot2",      # for awesome graphics
  "gridExtra",    # for grid.arrange() function
  "leaps",        # for regsubsets() function
  "plotly",       # for interactive plots
  "SMPracticals"  # for cement data
)

# Install required (CRAN) packages
for (pkg in pkgs) {
  if (!(pkg %in% installed.packages()[, "Package"])) {
    install.packages(pkg)
  }
}
```

]


---
class: clear, center, middle

```{r lets-go, echo=FALSE, out.width="70%"}
set.seed(4); RBitmoji::plot_comic(my_id, tag = "lets go")
```


---

# Removing the additive assumption

.font110[

* In our previous analyses of the Boston housing data, we assumed that the effect on median home value (`cmedv`) of increasing the average number of rooms per dwelling (`rm`) was independent of the percentage of lower status of the population (`lstat`)

* For example, the linear model $$E\left(cmedv\right) = \beta_0 + \beta_1 lstat + \beta_2 rm$$ states that the average effect on median home value of a one-unit increase in `rm` is always $\beta_2$, regardless of the percentage of lower status of the population

* By adding a *two-way interaction effect* between `rm` and `lstat`, we allow the effect of `rm` on `cmedv` to vary with `lstat` (and vice versa): $$E\left(cmedv\right) = \beta_0 + \beta_1 lstat + \beta_2 rm + 
\beta_3 lstat \times rm$$ 

]


---
class: clear, middle, center

```{r boston-plotly-01, echo=FALSE, out.width="100%"}
# Load required packages
library(plotly)

# Draw (interactive) 3-D scatterplot w/ fitted regression plane
plot_ly(data = pdp::boston, x = ~lstat, y = ~rm, z = ~cmedv, 
        mode = "markers", type = "scatter3d",
        marker = list(opacity = 0.3, symbol = 1, 
                      size = 5, color = "black")) %>%
  layout(
    scene = list(
      aspectmode = "manual", 
      aspectratio = list(x = 1, y = 1, z = 1)
    )
  )
```


---
class: clear, middle, center

```{r boston-plotly-02, echo=FALSE, out.width="100%"}
# Draw (interactive) 3-D scatterplot w/ fitted regression plane
fit <- lm(cmedv ~ lstat + rm, data = pdp::boston)
betas <- coef(fit)                
lstat <- seq(from = 1.73, to = 37.97, length = 50)
rm <- seq(from = 3.561, to = 8.780, length = 50)
yhat <- t(outer(lstat, rm, function(x1, x2) {
  betas[1] + betas[2]*x1 + betas[3]*x2
}))
plot_ly(x = ~lstat, y = ~rm, z = ~yhat, 
        type = "surface", opacity = 0.7) %>%
  add_trace(data = pdp::boston, x = ~lstat, y = ~rm, z = ~cmedv, 
            mode = "markers", 
            type = "scatter3d",
            marker = list(opacity = 0.7, symbol = 1, 
                          size = 5, color = "black")) %>%
  layout(
    scene = list(
      aspectmode = "manual", 
      aspectratio = list(x = 1, y = 1, z = 1),
      zaxis = list(title = "cmedv")
    )
  )
```


---
class: clear, middle, center

```{r boston-plotly-03, echo=FALSE, out.width="100%"}
# Draw (interactive) 3-D scatterplot w/ fitted regression plane (curvature)
fit2 <- lm(cmedv ~ lstat * rm, data = pdp::boston)
betas2 <- coef(fit2)  
yhat2 <- t(outer(lstat, rm, function(x1, x2) {
  betas2[1] + betas2[2]*x1 + betas2[3]*x2 + betas2[4]*x1*x2
}))
plot_ly(x = ~lstat, y = ~rm, z = ~yhat2, 
        type = "surface", opacity = 0.7) %>%
  add_trace(data = pdp::boston, x = ~lstat, y = ~rm, z = ~cmedv, 
            mode = "markers", 
            type = "scatter3d",
            marker = list(opacity = 0.7, symbol = 1, 
                          size = 5, color = "black")) %>%
  layout(
    scene = list(
      aspectmode = "manual", 
      aspectratio = list(x = 1, y = 1, z = 1),
      zaxis = list(title = "cmedv")
    )
  )

```


---

# Model-Building Problem 

.font120[

* "Conflicting" goals in regression model building:

    - Want **as many (useful) predictors as possible** so that the "information content" in the features will influence $\widehat{Y}$      

    - Want **as few predictors as necessary** because $Var\left(\widehat{Y}\right)$ increases with the number of predictors; also, sometimes collecting data on more features bd costly in terms of data collection/model maintenance

* Need to find a compromise that leads to the "best" regression equation

.center.blue[A .bold[parsimonious model] is the simplest model with the least assumptions and variables, but with the greatest explanatory power!]

]


---

# Some notes on variable selection techniques 

.font120[

* None of the variable selection techniques guarantees to find the "best" regression equation for the dataset of interest

* Different variable selection techniques may very well give different results

* Complete reliance on the algorithm for results is to be avoided

* Other valuable information such as experience with and knowledge of the data and problem should be utilized whenever possible

]


---

# Consequences of deleting variables

.font130[

* Improves the precision of the parameter estimates of retained variables

* Improves the precision of the variance of the predicted response

* Can induce bias into the estimates of coefficients and variance of the predicted response

]


---

# Some important questions

.font150[

1. Is at least one of the predictors $X_1, X_2, \dots, X_p$ useful in
predicting the response?

2. Do all the predictors help to explain $Y$ , or is only a subset
of the predictors useful?

3. How well does the model fit the data?

4. Given a set of predictor values, what response value should we predict, and how accurate is our prediction?

]


---

# Is at least one predictor useful?

.font150[

For the first question, we can use the general *F*-statistic $$F_{obs} = \frac{SSE(R) - SSE(F)}{df_R - df_F} \div \frac{SSE(F)}{df_F} = \frac{MSR}{MSE}$$

]

```{r, echo=FALSE}
fit <- lm(cmedv ~ ., data = pdp::boston)
```

.pull-left[

.font200.right.green[Boston housing example (all 15 predictors):]

]

.font140.pull-right[

| Quantity | Value |
|:---|:---|
| $RMSE$ | `r round(sigma(fit), digits = 3)` |
| $R^2$ | `r round(summary(fit)$r.squared, digits = 3)` |
| $R_{adj}^2$ | `r round(summary(fit)$adj.r.squared, digits = 3)` |
| $F_{obs}$ | `r round(summary(fit)$fstatistic, digits = 3)` |

]


---

# Deciding on the important variables

.font125[

* The most direct approach is called .bold[all subsets] or .bold[best subsets] regression: .purple[compute the least squares fit for all possible subsets and then choose between them based on some criterion that balances training error with model size.]

* It is often impossible to examine all possible models; for example, with 40 potential predictors, there are .bold[over a billion models!] Instead we often rely on automated approaches that search through a subset of them. We'll discuss a common approach called [stepwise regression](https://en.wikipedia.org/wiki/Stepwise_regression).

]

.font150.bold.center[

In general, with $p - 1$ predictors, there are $2^{p - 1}$ possible models!

]


---

# Hald cement data

.code125[

```{r cement-load}
# Load the Hald cement data
data(cement, package = "SMPracticals")
head(cement)  # see ?cement for details
```

]


---
class: clear, middle

.code130[

```{r cement-leaps-01}
# Load required packages
library(leaps)

# All subsets regression (main effects only)
a1 <- regsubsets(y ~ ., data = cement, 
                 nbest = 15, nvmax = 100)
```

]


---
class: clear, middle

```{r cement-leaps-02, echo=FALSE, fig.width=6, fig.asp=1, out.width="60%"}
# Plot results from all subsets regression
plot(a1, scale = "bic")
```


---
class: clear, middle

```{r cement-leaps-03, echo=FALSE, fig.width=6, fig.asp=0.618, out.width="100%"}
# Load required packages
library(ggplot2)

# Gather results
res1 <- data.frame(
  "nvar" = apply(summary(a1)$which, 1, FUN = function(x) sum(x) - 1),
  "bic" = summary(a1)$bic,
  "adjr2" = summary(a1)$adjr2
)

# Plot results
p1 <- ggplot(res1, aes(x = nvar, y = bic)) +
  geom_point(alpha = 0.5, size = 2, color = "darkred") +
  stat_summary(fun.y = min, geom = "line", alpha = 0.5, linetype = "dashed") +
  theme_light() +
  labs(x = "Number of predictors", y = "BIC")
p2 <- ggplot(res1, aes(x = nvar, y = adjr2)) +
  geom_point(alpha = 0.5, size = 2, color = "darkgreen") +
  stat_summary(fun.y = max, geom = "line", alpha = 0.5, linetype = "dashed") +
  theme_light() +
  labs(x = "Number of predictors", y = "Adjusted R-squared")
gridExtra::grid.arrange(p1, p2, nrow = 2)
```


---
class: clear, middle

.code100[

```{r cement-best-01, echo=FALSE}
# Summarize best model
summary(best1 <- lm(y ~ x1 + x2, data = cement))
```

]


---
class: clear, middle, center

```{r cement-best-02, echo=FALSE, fig.width=6, fig.asp=0.618, out.width="100%"}
# Plot residuals from best model
par(mfrow = c(1, 2))
plot(best1, which = 1:2)
```


---
class: clear, middle, center

.font300[

What about interactions?

]


---
class: clear, middle

.code130[

```{r cement-leaps-04}
# All subsets regression (with two-way interactions)
a2 <- regsubsets(y ~ .^2, data = cement, 
                 nbest = 40, nvmax = 1000)
```

]


---
class: clear, middle

```{r cement-leaps-05, echo=FALSE, fig.width=6, fig.asp=0.618, out.width="100%"}
# Gather results
res2 <- data.frame(
  "nvar" = apply(summary(a2)$which, 1, FUN = function(x) sum(x) - 1),
  "bic" = summary(a2)$bic,
  "adjr2" = summary(a2)$adjr2
)

# Plot results
p3 <- ggplot(res2, aes(x = nvar, y = bic)) +
  geom_point(alpha = 0.5, size = 2, color = "darkred") +
  stat_summary(fun.y = min, geom = "line", alpha = 0.5, linetype = "dashed") +
  scale_x_continuous(breaks = 1:10) +
  theme_light() +
  labs(x = "Number of predictors", y = "BIC")
p4 <- ggplot(res2, aes(x = nvar, y = adjr2)) +
  geom_point(alpha = 0.5, size = 2, color = "darkgreen") +
  stat_summary(fun.y = max, geom = "line", alpha = 0.5, linetype = "dashed") +
  scale_x_continuous(breaks = 1:10) +
  theme_light() +
  labs(x = "Number of predictors", y = "Adjusted R-squared")
gridExtra::grid.arrange(p3, p4, nrow = 2)
```


---
class: clear, middle

.code120[

```{r cement-leaps-06, echo=FALSE}
# Summarize best model
id <- which.min(summary(a2)$bic)
trms <- names(which(summary(a2)$which[id, ])[-1L])
form <- as.formula(paste("y ~", paste(trms, collapse = "+")))
round(summary(best2 <- lm(form, data = cement))$coefficients, digits = 3)
```

]


---
class: clear, middle

.font200.bold[The hierarchy principle:]

.font150.content-box-yellow[

If we include an interaction in a model, we should also include the main effects, even if the *p*-values associated with their coefficients are not significant

]

.font200.blue.center.bold[Why?]


---
class: clear, middle

.font300.bold[Pop quiz:]

.font200[

What is the maximum number of terms we could have if we allowed for three-way interaction effects (e.g., $x_1 x_2 x_3$)?

]


---
class: clear, middle, center

.font200[

Rather than search through all possible subsets (which becomes infeasible
for $p$ much larger than 40), we can seek a good path through them!

]


---

# Forward selection

.font125[

1. Begin with the **null model** - a model that contains an intercept, but no predictors

2. Fit $p - 1$ simple linear regressions and add to the null model the predictor that gives the .bold["biggest improvement"]

3. Add to that model the predictor that results in the .bold["biggest improvement"] amongst all two-predictor models

4. Continue until some stopping rule is satisfied, for example when all remaining variables have a *p*-value above some threshold

]

.red.font125.center[

Can still be used when $n \le p - 1$ (i.e., wide data)

]


---

# Backward selection

.font125[

1. Start with all candidate predictors in the model (.red[including interactions])

2. Fit $p - 1$ simple linear regressions and remove from the model the predictor that has the .bold["least impact on the fit"]


3. The new $\left(p - 1\right)$-predictor model is fit, and the predictor with the .bold["least impact on the fit"] is removed

4. Continue until some stopping rule is satisfied, for example when all remaining variables have a *p*-value above some threshold

]

.red.font125.center[

Requires that $n > p - 1$

]


---

# Selecting an "optimal" model

.font130[

There are a number of criteria for helping to select an "optimal" member in the path of models produced by forward or backward stepwise selection **in OLS**, for example:

* *Akaike information criterion*: $AIC = n\ln\left(SSE/n\right) + 2p$

* *Bayesian information criterion*: $BIC = n\ln\left(SSE/n\right) + \ln\left(n\right)p$

* Mallow's $C_p$; $C_p = AIC$ for linear regression with normal errors

* $R_{adj}^2 = 1 - \left(\frac{n - 1}{n - p}\right)\frac{SSE}{SST} = 1 - \left(\frac{n - 1}{n - p}\right)\left(1 - R^2\right)$

* $PRESS = \sum_{i = 1}^n e_{\left(i\right)}^2 = \sum_{i = 1}^n\left[e_i / \left(1 - h_i\right)\right]^2$

]


---

# Model selection in R

.font175[

* Base R's `step()` function and **MASS**'s `stepAIC()` function can be used to choose a model by AIC in a stepwise algorithm (i.e., forward, backward, or both)

* **leaps**'s `regsubsets()` function can be used to choose a model using an exhaustive search (i.e., best subsets), a stepwise algorithm (i.e., forward, backward, or both), or sequential replacement

]


---
class: clear, middle

.code100[

```{r boston-load}
# Load the Boston housing data
data(boston, package = "pdp")

# Print first few observations
head(tibble::as_tibble(boston), n = 5)
```

]


---
class: clear middle

.code115[

```{r press-function}
# Function to compute the PRESS statistic (a form of 
# cross-validation). Note: smaller is better!
PRESS <- function(object, ...) {
  if(!missing(...)) {
    res <- sapply(list(object, ...), FUN = function(x) {
      sum(rstandard(x, type = "predictive") ^ 2)
    })
    names(res) <- as.character(match.call()[-1L])
    res
  } else {
    sum(rstandard(object, type = "predictive") ^ 2)
  }
}
```

]


---
class: clear middle

.scrollable.code115[

```{r modelMetrics-function}
# Function to compute various model metrics
modelMetrics <- function(object, ...) {
  if(!missing(...)) {
    res <- sapply(list(object, ...), FUN = function(x) {
      c("AIC" = AIC(x), "BIC" = BIC(x), 
        "adjR2" = summary(x)$adj.r.squared,
        "RMSE"  = sigma(x), "PRESS" = PRESS(x), 
        "nterms" = length(coef(x)))
    })
    colnames(res) <- as.character(match.call()[-1L])
    res
  } else {
    c("AIC" = AIC(object), "BIC" = BIC(object), 
      "adjR2" = summary(object)$adj.r.squared, 
      "RMSE"  = sigma(object), "PRESS" = PRESS(object),
      "nterms" = length(coef(object)))
  }
}
```

]


---
class: clear, middle

.code115[

```{r boston-be}
# Backward elimination --------------------------

# Note that setting `k = 2` in the call to step(), which 
# is the default, corresponds to using AIC; below we set 
# it to `k = ln(n)`, which corresponds to using BIC!

# Main effects only (i.e., no interactions)
fit_max_1 <- lm(cmedv ~ ., data = boston)  #<<
be_1 <- step(fit_max_1, direction = "backward", 
             trace = 0, k = log(nrow(boston)))

# Main effects and two-way interactions
fit_max_2 <- lm(cmedv ~ .^2, data = boston)  #<<
be_2 <- step(fit_max_2, direction = "backward", 
             trace = 0, k = log(nrow(boston)))
```

]


---
class: clear, middle

.code115[

```{r boston-fs}
# Forward selection -----------------------------

# Main effects only (i.e., no interactions)
fit_min <- lm(cmedv ~ 1, data = boston)
fs_1 <- step(fit_min, direction = "forward", 
             scope = list(lower = fit_min,     #<<
                          upper = fit_max_1),  #<<
             trace = 0, k = log(nrow(boston)))

# Main effects and two-way interactions
fs_2 <- step(fit_min, direction = "forward", 
             scope = list(lower = fit_min,     #<<
                          upper = fit_max_2),  #<<
             trace = 0, k = log(nrow(boston)))
```

]


---
class: clear, middle

.code115[

```{r boston-ss}
# Stepwise selection ----------------------------

# Main effects only (i.e., no interactions)
ss_1 <- step(be_1, direction = "both", 
             scope = list(lower = fit_min,     #<<
                          upper = fit_max_1),  #<<
             trace = 0, k = log(nrow(boston)))

# Main effects and two-way interactions
ss_2 <- step(be_2, direction = "both", 
           scope = list(lower = fit_min,     #<<
                        upper = fit_max_2),  #<<
           trace = 0, k = log(nrow(boston)))
```

]


---
class: clear, middle

.code100[

```{r boston-compare-models, highlight.output = 6:7}
# Compare models
res <- modelMetrics(be_1, be_2, fs_1, fs_2, ss_1, ss_2)
round(res, digits = 3)
```

]


---
class: clear, middle

.scrollable[

```{r boston-ss-2}
summary(ss_2)
```

]


---
class: clear, middle

.scrollable[

```{r boston-fs-2}
summary(fs_2)
```

]


---

# Modern alternatives

.font130[

* [Regularized regression](https://koalaverse.github.io/AnalyticsSummit18/03-Regularization.html#1); for example, the LASSO

* Multivariate adaptive regression splines ([MARS](https://koalaverse.github.io/AnalyticsSummit18/04-MARS.html#1)): piecewise linear splines approach to multiple linear regression that automatically models nonlinearities and interactions between variables

  - Variable selection `r emo::ji("check")`

  - Nonlinear relationships `r emo::ji("check")`
    
  - Variable interactions `r emo::ji("check")`
    
  - Variable importance `r emo::ji("check")`

]


---
class: clear, middle, center

```{r quittin-time, echo=FALSE, out.width="60%"}
RBitmoji::plot_comic(my_id, tag = "quittin")
```
