---
title: "About me"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Who is this guy?

```{r bitmoji}
set.seed(101)
id <- "d47c6f08-0f4b-431c-b9aa-382377158c14-v1"
RBitmoji::plot_comic(id, tag = "math")
```


## Education

* B.S. in Statistics from WSU

* M.S. in Applied Statistics from WSU

* Ph.D. in Applied Mathematics from AFIT 
    
    * *Topics in Statistical Calibration*: http://www.dtic.mil/docs/citations/ADA598921
    

## Research interests

* Statistical/machines learning

* Interpretable machine learning (e.g., quantifying and visualizing both feature importance and feature effects)

* Statistical computing (e.g., developing useful/useless R packages)

* Inverse estimation (e.g., inverse prediction---maybe in this class? `r emo::ji("thinking")`)


## R packages

* [investr](https://cran.r-project.org/package=investr)

* [pdp](https://cran.r-project.org/package=pdp) 

* [vip](https://cran.r-project.org/package=vip)

* [sure](https://cran.r-project.org/package=sure)

* [RBitmoji](https://cran.r-project.org/package=RBitmoji)

* [roundhouse](https://cran.r-project.org/package=roundhouse)
